---
title: "DEPOXY Project — Physiological Analysis of Running Shoe Conditions"
subtitle: "Master STAPS IEAP | Python-R-Git Course"
author: "Baptiste AUGROS"
date: "`r format(Sys.Date(), '%B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    number_sections: true
    df_print: kable
    fig_caption: true
    code_folding: hide
    self_contained: true
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_file = file.path("results", "main.html"),
      ...
    )
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo       = TRUE,
  warning    = FALSE,
  message    = FALSE,
  fig.width  = 10,
  fig.height = 5.5,
  dpi        = 180,
  out.width  = "100%"
)

# ── Auto-install and load required packages ───────────────────────────────
required_pkgs <- c(
  "tidyverse", "ggplot2", "lme4", "lmerTest",
  "emmeans", "broom", "broom.mixed", "knitr",
  "kableExtra", "scales", "patchwork", "viridis"
)
for (pkg in required_pkgs) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    install.packages(pkg, repos = "https://cloud.r-project.org")
  }
  library(pkg, character.only = TRUE)
}

# ── DEPOXY colour palette ─────────────────────────────────────────────────
PALETTE <- c(A = "#E63946", B = "#2A9D8F", C = "#F4A261")

# ── Consistent kable helper ───────────────────────────────────────────────
nice_kable <- function(df, caption = "") {
  knitr::kable(df, caption = caption, digits = 3, align = "c") |>
    kableExtra::kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"),
      full_width = FALSE, position = "left"
    )
}

# ── Path helpers ──────────────────────────────────────────────────────────
# All paths are RELATIVE to the project root so the Rmd is portable.
DATA_DIR    <- file.path("data")
RESULTS_DIR <- file.path("results")
dir.create(RESULTS_DIR, showWarnings = FALSE, recursive = TRUE)

# ── Utility: read a results CSV silently ──────────────────────────────────
read_results <- function(fname) {
  path <- file.path(RESULTS_DIR, fname)
  if (!file.exists(path)) {
    warning(paste("File not found:", path,
                  "— run main.ipynb first."))
    return(NULL)
  }
  readr::read_csv(path, show_col_types = FALSE)
}
```

---

# Scientific Context {.unnumbered}

The **DEPOXY** project investigates peripheral energy expenditure during
treadmill running under three shoe conditions:

| Code | Shoe | Blocks |
|------|------|--------|
| **A** | KipRun KipStorm Foam | 1, 6 |
| **B** | KipRun KipStorm Carbon | 2, 5 |
| **C** | Prototype (non-commercialized) | 3, 4 |

The **ABCCBA counterbalanced design** controls for order and fatigue effects by
mirroring the shoe sequence. Six 5-minute running blocks are separated by
5-minute recoveries. Each condition is repeated twice; condition-level estimates
are the mean of the two trials.

**Primary outcomes:** VO₂, FC (heart rate), TSI/SmO₂, HHb, O₂Hb, and vertical
ground reaction force.

---

# Data Import {#data-import}

```{r load-merged}
# ── PRIMARY DATA SOURCE ───────────────────────────────────────────────────
# all_data_merged.csv is the single authoritative dataset produced by the
# Python pipeline (main.ipynb, Section 14).  It contains:
#   • Block-level physiological metrics (VO2, HR, TSI, HHb, AUC, slopes …)
#   • Participant metadata (participant_id, session_date, shoe_order,
#     block speeds, RPE values, and any future columns such as ATT)
#
# If the file does not exist yet, a synthetic placeholder is generated so
# the Rmd knits without error during template validation.
# ---------------------------------------------------------------------------

merged_path <- file.path(DATA_DIR, "all_data_merged.csv")

if (file.exists(merged_path)) {
  block_df <- readr::read_csv(merged_path, show_col_types = FALSE) |>
    mutate(condition = factor(condition, levels = c("A", "B", "C")))
  cat(sprintf("✓ all_data_merged.csv loaded: %d rows × %d columns\n",
              nrow(block_df), ncol(block_df)))
} else {
  # ── synthetic placeholder (run main.ipynb to generate the real file) ───
  warning("all_data_merged.csv not found — using synthetic placeholder data.")
  set.seed(42)
  block_df <- tibble(
    block           = 1:6,
    condition       = factor(c("A","B","C","C","B","A"), levels=c("A","B","C")),
    shoe_name       = c("KipRun Foam","KipRun Carbon","Prototype",
                        "Prototype","KipRun Carbon","KipRun Foam"),
    VO2_mean        = c(1000,1040,880,870,1035,1010) + rnorm(6, 0, 20),
    HR_mean         = c(152,158,145,148,156,150)     + rnorm(6, 0, 3),
    TSI_mean        = c(58,62,61,60,63,57)           + rnorm(6, 0, 1.5),
    HHb_mean        = c(-3,-1,0.3,0.2,-1.2,-2.8)    + rnorm(6, 0, 0.3),
    O2Hb_mean       = c(12,14,13,12.5,14.5,11.5)    + rnorm(6, 0, 0.5),
    Force_mean      = c(850,870,860,855,865,845)     + rnorm(6, 0, 10),
    VO2_auc         = VO2_mean * 150,
    HR_auc          = HR_mean  * 150,
    TSI_auc         = TSI_mean * 150,
    HHb_auc         = HHb_mean * 150,
    TSI_slope       = c(-0.002,-0.001,-0.003,-0.0025,-0.0015,-0.002),
    HHb_slope       = c( 0.001, 0.001, 0.002, 0.002,  0.001,  0.001),
    FC_drift        = c( 0.8,   1.1,   0.6,   0.7,    1.0,    0.9),
    step_freq_bpm   = c(168,170,167,168,171,169),
    n_steps         = c(840,850,835,840,855,845),
    # metadata columns (placeholder values)
    participant_id  = "LEPLAT_Allan",
    session_date    = as.Date("2026-02-20"),
    shoe_order      = "ABCCBA",
    speed           = c(12,12,12,12,12,12),
    rpe             = c(12,13,11,11,13,12)
  )
}

# ── Load statistical output tables (produced by Python pipeline) ──────────
anova_df     <- read_results("anova_results.csv")
posthoc_df   <- read_results("posthoc_bonferroni.csv")
friedman_df  <- read_results("friedman_test.csv")
corr_df      <- read_results("pairwise_correlations.csv")
corr_mat_raw <- read_results("correlation_matrix.csv")
normality_df <- read_results("normality_tests.csv")
baseline_df  <- read_results("baseline_comparison.csv")

# Reconstruct condition-level summary from block_df (no need for separate CSV)
condition_df <- block_df |>
  group_by(condition) |>
  summarise(
    across(where(is.numeric), list(mean = ~mean(., na.rm = TRUE),
                                   sd   = ~sd(.,   na.rm = TRUE)),
           .names = "{.col}_{.fn}"),
    .groups = "drop"
  )
```

---

# Participant Metadata {#metadata}

```{r metadata-table}
# Display all metadata columns present in the merged file.
# New columns (e.g. adipose_tissue_thickness) appear automatically.
meta_cols <- c("participant_id", "session_date", "shoe_order",
               intersect(names(block_df),
                         c("speed", "rpe", "adipose_tissue_thickness",
                           # add any future column names here or leave
                           # the intersect to catch them automatically:
                           grep("^(?!block|condition|shoe|VO2|HR|TSI|HHb|O2|Force|auc|slope|drift|step|n_)",
                                names(block_df), value = TRUE, perl = TRUE))))

meta_display <- block_df |>
  distinct(across(any_of(meta_cols)))

if (ncol(meta_display) > 0 && nrow(meta_display) > 0) {
  meta_display |>
    nice_kable("Participant metadata (from all_data_merged.csv)")
} else {
  cat("No participant-level metadata columns detected in all_data_merged.csv.")
}
```

```{r block-metadata-table}
# Block-specific metadata: speed and RPE per block (and any future block vars)
block_meta_cols <- intersect(names(block_df),
                              c("block", "condition", "shoe_name",
                                "speed", "rpe", "adipose_tissue_thickness"))
if (length(block_meta_cols) > 2) {
  block_df |>
    select(all_of(block_meta_cols)) |>
    nice_kable("Block-level metadata (speed, RPE, …)")
}
```

---

# Descriptive Statistics {#descriptive}

## Block-level overview

```{r block-table}
block_df |>
  select(block, condition, shoe_name,
         any_of(c("VO2_mean","HR_mean","TSI_mean","HHb_mean","Force_mean"))) |>
  rename_with(~ case_match(.,
    "block"      ~ "Block",
    "condition"  ~ "Condition",
    "shoe_name"  ~ "Shoe",
    "VO2_mean"   ~ "VO₂ (mL/min)",
    "HR_mean"    ~ "FC (bpm)",
    "TSI_mean"   ~ "TSI (%)",
    "HHb_mean"   ~ "HHb (µM)",
    "Force_mean" ~ "Peak Force (N)",
    .default = .
  )) |>
  nice_kable("Block-level steady-state means (primary outcomes)")
```

## Condition-level summary (A / B / C)

```{r condition-summary}
block_df |>
  group_by(condition) |>
  summarise(
    across(any_of(c("VO2_mean","HR_mean","TSI_mean","HHb_mean","Force_mean")),
           list(mean = ~mean(., na.rm = TRUE),
                sd   = ~sd(.,   na.rm = TRUE)),
           .names = "{.col}_{.fn}"),
    .groups = "drop"
  ) |>
  nice_kable("Condition-level means ± SD (averaged across two trials)")
```

---

# Visualisations {#viz}

## Primary Outcomes by Condition

```{r fig-boxplots, fig.height=9, fig.cap="Physiological responses by shoe condition. Each point = one running block trial."}
vars_info <- list(
  list(col = "VO2_mean",   label = "VO₂ (mL/min)"),
  list(col = "HR_mean",    label = "FC / HR (bpm)"),
  list(col = "TSI_mean",   label = "TSI / SmO₂ (%)"),
  list(col = "HHb_mean",   label = "HHb (µM)"),
  list(col = "O2Hb_mean",  label = "O₂Hb (µM)"),
  list(col = "Force_mean", label = "Peak Force (N)")
)

bp_plots <- lapply(vars_info, function(v) {
  if (!v$col %in% names(block_df)) return(NULL)
  ggplot(block_df, aes(x = condition, y = .data[[v$col]],
                        fill = condition, colour = condition)) +
    geom_boxplot(alpha = 0.35, outlier.shape = NA,
                 width = 0.5, linewidth = 0.7) +
    geom_jitter(width = 0.1, size = 3.5, alpha = 0.85) +
    scale_fill_manual(values = PALETTE)   +
    scale_colour_manual(values = PALETTE) +
    labs(title = v$label, x = "Condition", y = v$label) +
    theme_minimal(base_size = 11) +
    theme(legend.position = "none",
          plot.title = element_text(face = "bold", size = 10))
})
bp_plots <- Filter(Negate(is.null), bp_plots)
p_box <- patchwork::wrap_plots(bp_plots, ncol = 3)
p_box
ggsave(file.path(RESULTS_DIR, "R_Fig1_Condition_Boxplots.png"),
       p_box, width = 14, height = 9, dpi = 180)
```

## Area Under Curve by Condition

```{r fig-auc, fig.height=4, fig.cap="Mean AUC ± SD per condition."}
auc_vars <- intersect(c("VO2_auc","HR_auc","TSI_auc","HHb_auc"), names(block_df))
auc_labels <- c(VO2_auc="VO₂ AUC", HR_auc="FC AUC",
                TSI_auc="TSI AUC", HHb_auc="HHb AUC")

p_auc <- block_df |>
  select(condition, all_of(auc_vars)) |>
  pivot_longer(-condition, names_to = "Variable", values_to = "AUC") |>
  mutate(Variable = recode(Variable, !!!auc_labels)) |>
  group_by(condition, Variable) |>
  summarise(mean = mean(AUC, na.rm = TRUE),
            sd   = sd(AUC,   na.rm = TRUE),
            .groups = "drop") |>
  ggplot(aes(x = condition, y = mean, fill = condition)) +
    geom_col(colour = "black", linewidth = 0.4, alpha = 0.85) +
    geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0.25) +
    scale_fill_manual(values = PALETTE) +
    facet_wrap(~Variable, scales = "free_y", nrow = 1) +
    labs(title = "Area Under Curve by Condition",
         x = "Condition", y = "AUC") +
    theme_minimal(base_size = 11) +
    theme(legend.position = "none",
          strip.text = element_text(face = "bold"))
p_auc
ggsave(file.path(RESULTS_DIR, "R_Fig2_AUC_By_Condition.png"),
       p_auc, width = 12, height = 4, dpi = 180)
```

## Signal Slopes by Condition

```{r fig-slopes, fig.height=4, fig.cap="Mean linear slopes ± SD. FC drift in bpm/min; TSI and HHb slopes in units/s."}
slope_vars <- intersect(c("TSI_slope","HHb_slope","FC_drift"), names(block_df))
slope_labels <- c(TSI_slope = "TSI slope (units/s)",
                  HHb_slope = "HHb slope (units/s)",
                  FC_drift  = "FC drift (bpm/min)")

p_slopes <- block_df |>
  select(condition, all_of(slope_vars)) |>
  pivot_longer(-condition, names_to = "Variable", values_to = "slope") |>
  mutate(Variable = recode(Variable, !!!slope_labels)) |>
  group_by(condition, Variable) |>
  summarise(mean = mean(slope, na.rm = TRUE),
            sd   = sd(slope,   na.rm = TRUE),
            .groups = "drop") |>
  ggplot(aes(x = condition, y = mean, fill = condition)) +
    geom_col(colour = "black", linewidth = 0.4, alpha = 0.85) +
    geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0.25) +
    geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40") +
    scale_fill_manual(values = PALETTE) +
    facet_wrap(~Variable, scales = "free_y", nrow = 1) +
    labs(title = "Signal Slopes by Shoe Condition",
         x = "Condition", y = "Slope") +
    theme_minimal(base_size = 11) +
    theme(legend.position = "none",
          strip.text = element_text(face = "bold"))
p_slopes
ggsave(file.path(RESULTS_DIR, "R_Fig3_Slopes_By_Condition.png"),
       p_slopes, width = 12, height = 4, dpi = 180)
```

## Baseline Recovery

```{r fig-baseline, fig.height=4, fig.cap="Initial vs final baseline per metric. Δ = Final − Initial."}
if (!is.null(baseline_df)) {
  p_bl <- baseline_df |>
    pivot_longer(c(Baseline_Initial, Baseline_Final),
                 names_to = "Phase", values_to = "Value") |>
    mutate(Phase = recode(Phase,
                          Baseline_Initial = "Initial",
                          Baseline_Final   = "Final")) |>
    ggplot(aes(x = Phase, y = Value, fill = Phase)) +
      geom_col(colour = "black", linewidth = 0.4, alpha = 0.85, width = 0.5) +
      scale_fill_manual(values = c(Initial = "#455A64", Final = "#F57F17")) +
      geom_text(data = baseline_df,
                aes(x = 1.5,
                    y = pmax(Baseline_Initial, Baseline_Final, na.rm = TRUE) * 1.05,
                    label = sprintf("Δ = %+.2f", Delta_Final_minus_Initial)),
                inherit.aes = FALSE, size = 3.2, colour = "grey20") +
      facet_wrap(~Metric, scales = "free_y") +
      labs(title = "Baseline Comparison: Initial vs Final",
           x = "Phase", y = "Value") +
      theme_minimal(base_size = 11) +
      theme(legend.position = "none",
            strip.text = element_text(face = "bold"))
  p_bl
  ggsave(file.path(RESULTS_DIR, "R_Fig4_Baseline_Comparison.png"),
         p_bl, width = 12, height = 4, dpi = 180)
} else {
  cat("Baseline comparison data not available — run main.ipynb first.")
}
```

## Correlation Scatter Plots

```{r fig-corr-scatter, fig.height=7, fig.cap="Required pairwise scatter plots with linear regression line."}
pairs_to_plot <- list(
  list(x = "VO2_mean",  y = "TSI_mean",   xl = "VO₂ (mL/min)", yl = "TSI (%)"),
  list(x = "VO2_mean",  y = "HHb_mean",   xl = "VO₂ (mL/min)", yl = "HHb (µM)"),
  list(x = "VO2_mean",  y = "HR_mean",    xl = "VO₂ (mL/min)", yl = "FC (bpm)"),
  list(x = "HR_mean",   y = "Force_mean", xl = "FC (bpm)",      yl = "Peak Force (N)")
)

scatter_plots <- lapply(pairs_to_plot, function(p) {
  if (!all(c(p$x, p$y) %in% names(block_df))) return(NULL)
  r_val <- cor(block_df[[p$x]], block_df[[p$y]], use = "complete.obs")
  ggplot(block_df, aes(x = .data[[p$x]], y = .data[[p$y]], colour = condition)) +
    geom_point(size = 4, alpha = 0.85) +
    geom_smooth(method = "lm", se = TRUE, colour = "grey30",
                linewidth = 0.8, linetype = "dashed",
                inherit.aes = FALSE,
                aes(x = .data[[p$x]], y = .data[[p$y]])) +
    scale_colour_manual(values = PALETTE) +
    annotate("text", x = -Inf, y = Inf,
             label = sprintf("r = %.2f", r_val),
             hjust = -0.1, vjust = 1.5, size = 3.8,
             colour = "grey20", fontface = "italic") +
    labs(x = p$xl, y = p$yl, colour = "Condition") +
    theme_minimal(base_size = 11) +
    theme(legend.position = "bottom",
          legend.title = element_text(face = "bold"))
})
scatter_plots <- Filter(Negate(is.null), scatter_plots)
p_scatter <- patchwork::wrap_plots(scatter_plots, ncol = 2)
p_scatter
ggsave(file.path(RESULTS_DIR, "R_Fig5_Correlation_Scatter.png"),
       p_scatter, width = 12, height = 7, dpi = 180)
```

## Correlation Heatmap

```{r fig-corr-heatmap, fig.height=6, fig.cap="Full Pearson correlation matrix (block-level metrics)."}
if (!is.null(corr_mat_raw)) {
  cm_long <- corr_mat_raw |>
    tibble::rownames_to_column("Var1") |>
    pivot_longer(-Var1, names_to = "Var2", values_to = "r")

  p_heatmap <- ggplot(cm_long, aes(x = Var1, y = Var2, fill = r)) +
    geom_tile(colour = "white", linewidth = 0.5) +
    geom_text(aes(label = round(r, 2)), size = 2.8) +
    scale_fill_gradient2(low  = "#E63946", mid  = "white", high = "#2A9D8F",
                         midpoint = 0, limits = c(-1, 1), name = "r") +
    labs(title = "Correlation Matrix — Block-Level Metrics",
         x = NULL, y = NULL) +
    theme_minimal(base_size = 10) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          panel.grid  = element_blank())
  p_heatmap
  ggsave(file.path(RESULTS_DIR, "R_Fig6_Correlation_Heatmap.png"),
         p_heatmap, width = 10, height = 6, dpi = 180)
}
```

---

# Statistical Analyses {#stats}

## Normality Tests (Shapiro-Wilk)

```{r normality-table}
if (!is.null(normality_df)) {
  normality_df |>
    mutate(normal = ifelse(normal, "✓ Yes", "✗ No")) |>
    nice_kable("Shapiro-Wilk normality test (block-level observations)")
}
```

## Repeated-Measures ANOVA (lmerTest)

```{r lme4-anova}
# Artificial repeat structure: rep1 = first occurrence of each condition,
# rep2 = second occurrence (ABCCBA design has exactly 2 per condition).
condition_long <- block_df |>
  mutate(repeat_id = case_when(
    block %in% c(1, 2, 3) ~ "rep1",
    TRUE                   ~ "rep2"
  ))

lme_targets <- intersect(
  c("VO2_mean","HR_mean","TSI_mean","HHb_mean",
    "VO2_auc","HR_auc","TSI_slope","HHb_slope","FC_drift"),
  names(block_df)
)

lme_results <- lapply(lme_targets, function(var) {
  df_mod <- condition_long |>
    select(condition, repeat_id, value = all_of(var)) |>
    drop_na()
  if (nrow(df_mod) < 4) return(NULL)
  tryCatch({
    m <- lmerTest::lmer(value ~ condition + (1 | repeat_id), data = df_mod)
    broom.mixed::tidy(m) |>
      filter(effect == "fixed") |>
      mutate(variable = var, .before = 1) |>
      select(variable, term, estimate, std.error, statistic, p.value)
  }, error = function(e) NULL)
})

lme_table <- bind_rows(Filter(Negate(is.null), lme_results))
if (nrow(lme_table) > 0) {
  lme_table_out <- lme_table |>
    mutate(across(where(is.numeric), round, 4),
           p.value = ifelse(p.value < 0.001, "<0.001",
                            as.character(round(p.value, 4))))
  lme_table_out |>
    nice_kable("Linear mixed-effects model (lmerTest): condition effect estimates")
  readr::write_csv(lme_table, file.path(RESULTS_DIR, "R_lme_results.csv"))
}
```

## Post-hoc Comparisons — emmeans (Bonferroni)

```{r emmeans}
em_results <- lapply(lme_targets, function(var) {
  df_mod <- condition_long |>
    select(condition, repeat_id, value = all_of(var)) |>
    drop_na()
  if (nrow(df_mod) < 4) return(NULL)
  tryCatch({
    m  <- lmerTest::lmer(value ~ condition + (1 | repeat_id), data = df_mod)
    em <- emmeans::emmeans(m, "condition")
    emmeans::contrast(em, method = "pairwise", adjust = "bonferroni") |>
      broom::tidy() |>
      mutate(variable = var, .before = 1)
  }, error = function(e) NULL)
})

em_table <- bind_rows(Filter(Negate(is.null), em_results))
if (nrow(em_table) > 0) {
  em_out <- em_table |>
    select(variable, contrast, estimate, std.error, statistic, adj.p.value) |>
    mutate(across(where(is.numeric), round, 4),
           adj.p.value = ifelse(adj.p.value < 0.001, "<0.001",
                                as.character(round(adj.p.value, 4))))
  em_out |>
    nice_kable("Post-hoc pairwise comparisons (emmeans, Bonferroni correction)")
  readr::write_csv(em_table, file.path(RESULTS_DIR, "R_emmeans_posthoc.csv"))
}
```

## Python ANOVA Import (η² & Cohen's d)

```{r python-anova}
if (!is.null(anova_df) && nrow(anova_df) > 0) {
  anova_df |>
    select(any_of(c("Variable","F","p_value","eta_sq","significant"))) |>
    nice_kable("Repeated-measures ANOVA (Python statsmodels) with η²")
}
if (!is.null(posthoc_df) && nrow(posthoc_df) > 0) {
  posthoc_df |>
    select(any_of(c("Variable","Comparison","t_stat",
                    "p_bonferroni","cohens_d","significant"))) |>
    nice_kable("Post-hoc Bonferroni pairwise tests with Cohen's d (Python)")
}
```

## Friedman Non-parametric Test

```{r friedman}
if (!is.null(friedman_df) && nrow(friedman_df) > 0) {
  friedman_df |>
    mutate(significant = ifelse(significant, "Yes", "No")) |>
    nice_kable("Friedman test (non-parametric robustness check — Python)")
} else {
  # Fall back to R-side Friedman
  fr_results <- lapply(lme_targets, function(var) {
    wide <- block_df |>
      select(condition, value = all_of(var)) |>
      drop_na() |>
      group_by(condition) |>
      mutate(rep = row_number()) |>
      pivot_wider(names_from = condition, values_from = value)
    if (ncol(wide) < 4) return(NULL)
    mat <- wide |> select(-rep) |> as.matrix()
    tryCatch({
      ft <- friedman.test(mat)
      tibble(Variable   = var,
             statistic  = round(ft$statistic, 3),
             p_value    = round(ft$p.value, 4),
             significant = ft$p.value < 0.05)
    }, error = function(e) NULL)
  })
  fr_table <- bind_rows(Filter(Negate(is.null), fr_results))
  if (nrow(fr_table) > 0) {
    fr_table |>
      mutate(significant = ifelse(significant, "Yes", "No")) |>
      nice_kable("Friedman test (R-computed)")
    readr::write_csv(fr_table, file.path(RESULTS_DIR, "R_friedman_results.csv"))
  }
}
```

---

# Correlation Analysis {#correlations}

```{r corr-pairs}
if (!is.null(corr_df)) {
  corr_df |>
    mutate(interpretation = case_when(
      abs(r) >= 0.7 ~ "Strong",
      abs(r) >= 0.4 ~ "Moderate",
      TRUE          ~ "Weak"
    )) |>
    nice_kable("Pairwise Pearson correlations (block level, n = 6)")
}
```

---

# Discussion {#discussion}

## VO₂ and Running Economy

Differences in VO₂ between conditions at an identical running speed directly
reflect differences in running economy. Carbon-plate footwear (condition B) is
hypothesized to reduce metabolic cost through elastic energy storage and return;
a lower VO₂ for B relative to A and C would support this mechanism.

## Cardiovascular Response (FC / HR)

Heart rate is treated as a full physiological modality rather than a secondary
descriptor. The cardiac drift slope (bpm/min) captures cumulative cardiovascular
loading across the block — a steeper positive drift signals greater strain under
that shoe condition, which has implications for pacing and fatigue during
prolonged efforts.

## Muscle Oxygenation Dynamics

The TSI oxygenation slope quantifies the rate at which peripheral O₂ extraction
outpaces delivery. A steeper negative slope indicates faster deoxygenation under
that condition. O₂Hb (= HbTot − HHb) provides a complementary index of bound
oxygen availability. Recovery slopes (from the recovery blocks) indicate how
quickly the muscle re-oxygenates when running stops.

## Baseline Recovery

Comparing initial and final baselines tests whether cumulative fatigue has
altered resting physiology by the end of the protocol. Persistent VO₂ elevation
or TSI depression at the final baseline would indicate incomplete recovery
between Blocks 5–6 and the final rest, and should be flagged as a confounder.

## Forward-Compatibility Note

The pipeline is designed so that adding new variables — such as
adipose tissue thickness (ATT) for NIRS signal correction — requires no
code changes. Any new column in `Summary_file_DEPOXY_Project.csv` propagates
automatically through `load_participant_metadata()`, `merge_metadata_with_block_data()`,
and into `all_data_merged.csv`, after which it appears in this report's
metadata tables without modification.

## Statistical Limitations

With a single participant and n = 2 repetitions per condition, all inferential
statistics are exploratory. Effect sizes (η², Cohen's d) serve as inputs for
future sample-size planning rather than as confirmatory evidence. The Friedman
test provides a non-parametric sanity check that does not assume normality or
sphericity.

---

# Session Information {.unnumbered}

```{r session-info, echo=FALSE}
sessionInfo()
```

---

*Report generated from `main.Rmd` · data source: `data/all_data_merged.csv`*  
*DEPOXY Project | Baptiste AUGROS | Master STAPS IEAP 2026*
